Supplemental Materials for the paper: \\Pesticides pollution of small streams in Germany

### 2016-08-11 article.tex
add numbers to overview, finalized pollution, added s(x) + s(y) + ti(x, y) to test for interaction in GAM

We queried chemical monitoring data of pesticides from sampling sites with catchment size $\mathrm{< 100km^2}$ for the years 2005 to 2015 from all 13 non-city federal states of Germany.
Additionally, we compiled data available from previous studies and searched online databases.
This yielded to a total of more then 30 datasets of different formats.
In the following we will use the ISO 3166-2:DE standard abbreviations for federal states.

We homogenized and unified these datasets into a common database.
We implemented a robust and transparent data cleaning work flow \citep{poisot_best_2015}, though parts of the dataset are proprietary.
An overview of the data cleaning process is provided in the supplemental materials.  
To assess whether samples were taken during potential rainfall events we intersected sampling coordinates with daily precipitation data \citep{rauthe_central_2013} from the sampling date and the day before.

@@

The compiled dataset comprised only few standing waters (58 sites) and the majority (90\%) of samples where taken via grab sampling.  % see clean.R for numbers
Therefore, we report only results of grab samples from streams. 
The analysed dataset comprised 42236 samples from 3049 sampling sites.  %see do_overview.R for numbers.
We found big differences in the number of sampling sites between federal states (Figure \ref{fig:fig1} and Supplement, Table \todo{Set ref to table}).

@@

In total 484 different compounds that could be classified as pesticides and their metobalites were measured at least once (Supplement, Table \todo{Set ref to table}). 
Most of the compounds were herbicides (179), followed by insecticides (117) and fungicides (109).
We found substantial differences of the spectra of analysed compounds (Figure \ref{fig:figvar}).
Hierarchical clustering revealed three groups of states:
i) with less then 100 compounds (SL, ST and TH), ii) with a medium sized spectra and iii) with a big and distinct spectra (RP and NI).
Only 5\% of the samples were taken at or after days with rainfall events greater than 10mm / day.

\begin{figure}
  \includegraphics[width=\textwidth]{fig/figvar.pdf}
  \caption{Compound spectra of the different federal states. Left: Barcode plot - Each vertical line is an analysed compound. Right: MDS ordination. 
  Colors according to three groups determined by hierarchical clustering (see Supplement Figure xxx).}
  \label{fig:figvar}
\end{figure}
\todo{prettify figure! check cowplot package.}

We were able to derive for 2376 sites catchment sizes and the proportion of agriculture within catchments. \todo{Check why there are so many missing in BW, NW, RP, SN and TH = phchsitesinfo}.
The distribution of sampling sites across catchment area and agricultural area in the catchment revealed a sharp decline in the distribution of catchment-sizes below $10~km^2$, with most sampling sites with catchments between 10 and 25 $km^2$ (Figure \ref{fig:figezglu}).
The proportion of agriculture in the catchments decreased with increasing catchment size.

@@

\begin{figure}
  \includegraphics[width=0.8\textwidth]{fig/figrac.pdf}
  \caption{Effect of agriculture within the catchment (left) and catchment size (right) on the number of RAC exceedances. Red line marks statistically significant changes. Dashed lines denote 95\% Confidence Intervals.
  }
  \label{fig:figrac}
\end{figure}

\subsection{Pesticide pollution of small streams}

>>>>

We queried chemical monitoring data of pesticides from sampling sites with catchment sizes $\mathrm{< 100km^2}$ for the years 2005 to 2015 from all 13 non-city federal states of Germany.
Additionally, we compiled data available from previous studies and searched online databases.
This yielded to a total of more then 30 datasets of different formats.
In the following we will use the ISO 3166-2:DE standard abbreviations for federal states.

We homogenized and unified these datasets into a common database.
We implemented a robust and transparent data cleaning work flow \citep{poisot_best_2015}, though parts of the dataset are proprietary.
An overview of the data cleaning process is provided in the supplemental materials.  
To assess whether samples were taken during potential rainfall events we performed spatio-temporal intersection of sampling events with daily precipitation data from the sampling date and the day before. \citep{rauthe_central_2013}.

@@

The compiled dataset comprised only few standing waters (58 sites) and the majority (90\%) of samples where taken via grab sampling.  % see clean.R for numbers
Therefore, we report only results of grab samples from streams. 
The analysed dataset comprised 2,918,604 measurements from 42,236 samples from 3,049 sampling sites.  %see do_overview.R for numbers.
We found big differences in the number of sampling sites between federal states (Figure \ref{fig:fig1} and Supplement, Table \todo{Set ref to table}).

@@

In total 484 different compounds used as pesticides and their metobalites were measured at least once (Supplement, Table \todo{Set ref to table}). 
Most of the compounds were herbicides (179), followed by insecticides (117) and fungicides (109).
We found substantial differences of the spectra of analysed compounds between federal states (Figure \ref{fig:figvar}).
Hierarchical clustering revealed three groups:
i) with less then 100 compounds (SL, ST and TH), \\
ii) with a medium sized spectra and \\
iii) with a big and distinct spectra (RP and NI).

Only 5.5\% (160,800) of all measurements were detects above the limit of quantification (LOQ).
The spatio-temporal intersection revealed that 5\% of the samples were taken at or after days with rainfall events greater than 10mm / day. \todo{ref supplement}.

\begin{figure}
  \includegraphics[width=\textwidth]{fig/figvar.pdf}
  \caption{Compound spectra of the different federal states. Left: Barcode plot - Each vertical line is an analysed compound. Right: MDS ordination. 
  Colors according to three groups determined by hierarchical clustering (see Supplement Figure xxx).}
  \label{fig:figvar}
\end{figure}

We were able to derive for 2376 sites catchment sizes and the proportion of agriculture within catchments. 
The distribution of sampling sites across catchment area and agricultural area in the catchment revealed a sharp decline in the distribution of catchment-sizes below $10~km^2$, with most sampling sites with catchments between 10 and 25 $km^2$ (Figure \ref{fig:figezglu}).
The proportion of agriculture in the catchments decreased with increasing catchment size.

@@

Modeling the number of RAC exceedances as function of agriculture within catchment and catchment size revealed that there is a strong and statistically significant increase up to 25\% agriculture.
Above this threshold the exceedances level off followed by a increase above 75\% (Figure \ref{fig:figrac}, left).

We could no detect any effect of catchment size on the number of RAC exceedances (Figure \ref{fig:figrac}, right) and no interaction between these two predictors.

\begin{figure} [h]
  \includegraphics[width=0.8\textwidth]{fig/figrac.pdf}
  \caption{Effect of agriculture within the catchment (left) and catchment size (right) on the number of RAC exceedances. Red line marks statistically significant changes. Dashed lines denote 95\% Confidence Intervals.
  }
  \label{fig:figrac}
\end{figure}

The number of EQS-exceedances and the 95th percentile of $TU_{max}$ showed similar patterns and thresholds (see Supplement Figure Sxxx - Sxxx). \todo{references}.

\subsection{Pesticide pollution of small agricultural streams in germany}
Based on the results previous results and given the low amount of sample sites with catchment sizes below 10 km\textsuperscript{2} we explored the pollution of small agricultural streams, defined as streams with catchment size less than 30 km\textsuperscript{2} and more than 25\% agriculture within the catchment (n = 1030 sites).

% EQS
% RAC
% TU
% Mixture

### 2016-08-11 supplement/Supplemental_Materials.tex
add numbers to overview, finalized pollution, added s(x) + s(y) + ti(x, y) to test for interaction in GAM

\section{RAC Exceedances}
\begin{figure}[h]
	\centering
	\includegraphics[width = 0.6\textwidth]{ezgagrirac}
	\caption{Raw data used for the model in equation xxx and Figure xxx of the main article. Color codes the number of RAC exceedances (on a log-scale). Grey points are sites without any exceedance.}
	\label{fig:ezgagrirac}
\end{figure}
\todo{update refs}

@@

With $y_i$ being $log(TU_{max~i})$ and $x_i$ the predictor functions $f_1(Agri_i) + f_2(Size_i)$. $P(censored|x_i)$ is the probability of censoring at given predictors. Further, we assumed constant variance. This model was fitted using the gamlss package \citep{stasinopoulos_generalized_2007}.

@@

\pagebreak
\section{EQS Exceedances}
We modeled the number of EQS exceedances (No) the same way as RAC exceedances (see main article).

\begin{figure}[h]
	\centering
	\includegraphics[width = 0.6\textwidth]{ezgagrieqs}
	\caption{Raw data used for modelling EQS exceedances. Color codes the number of EQS exceedances (on a log-scale). Grey points are sites without any exceedance.}
	\label{fig:ezgagrieqs}
\end{figure}

>>>>

\section{EQS Exceedances}
We modeled the number of EQS exceedances (No) the same way as RAC exceedances (see main article).

\begin{figure}[h]
	\centering
	\includegraphics[width = 0.6\textwidth]{ezgagrieqs}
	\caption{Raw data used for modelling EQS exceedances. Color codes the number of EQS exceedances (on a log-scale). Grey points denote sites without any exceedance.}
	\label{fig:ezgagrieqs}
\end{figure}

\section{RAC Exceedances}
\begin{figure}[h]
	\centering
	\includegraphics[width = 0.6\textwidth]{ezgagrirac}
	\caption{Raw data used for the model in equation xxx and Figure xxx of the main article. Color codes the number of RAC exceedances (on a log-scale). Grey points denote sites without any exceedance.}
	\label{fig:ezgagrirac}
\end{figure}
\todo{update refs}

@@

With $y_i$ being $log(TU_{max~i})$ and $x_i$ the predictor functions $f_1(Agri_i) + f_2(Size_i)$. $P(censored|x_i)$ is the probability of censoring at given predictors and $E(y_i | y_i > -8.75))$ the expectet value given non-censoring. Further, we assumed constant variance. This model was fitted using the gamlss package \citep{stasinopoulos_generalized_2007}.

@@
